---
title: "Reproducible Research Project 1"
author: "Ismael Luna"
output: md_document
---

## Loading and preprocessing the data

#### Load the data (i.e. read.csv())
```{r}
activity <- read.csv("./activity.csv")
```

### Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
activity$date <- as.POSIXct(activity$date, format = "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)
summary(activity)
```


## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day
```{r}
total_step_activity <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
total_step_activity
```

### If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r}
names(total_step_activity) <- c("date", "steps")
hist(total_step_activity$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red", ylim = c(0,20), breaks = seq(0,25000, by = 2500))
```

### Calculate and report the mean and median of the total number of steps taken per day
```{r}
### Mean 
mean(total_step_activity$steps)
### Median 
median(total_step_activity$steps)
```

## What is the average daily activity pattern?

### Make a time series plot (i.e.type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
average_daily_activity <- aggregate(activity$steps, by = list(activity$interval), FUN = mean, na.rm = TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="red", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```

### Interval with maximun number of steps
```{r}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```


## Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity$steps))
```

### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r New Data set}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]

```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imputed_activity <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, imputed_activity, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

hist(total_steps_imputed$daily_steps, col = "red", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```


### Mean of the total number of steps taken per day
```{r}
mean(total_steps_imputed$daily_steps)
```

### Median of the total steps taken per day
```{r}
median(total_steps_imputed$daily_steps)
```

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
require(ggplot2)

activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "sábado" | weekdays(x) =="domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })

```

### Make a panel plot containing a time series plot (i.e.type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
